\TheTitle

### 2017-07-13 paper/discussion.tex
Cleaning up todos
Removing some TODO's, removing some hyphenations.
Trying to clarify discussion a bit

% vim:set et sw=2 ts=4 tw=72:
% Jun 24, 2017
% TODO: Go back to the conceptual task results and make them stronger

\dmg{don't editorialize: simply say: the results show that Linvis is superior to ... in ....
as we discussed today, it is about the tools more than the model
}

Overall, the results show fairly conclusive results; the DAG is unable
to adequately provide users with an explanation of the changes being
made in the repository, the DAG is not easily understood, and the DAG is
unable to provide summarizations of the changes being made at a merge.
Meanwhile, the \mt model is able to provide a better explanation of how
a commit is integrated, along with better summarization capabilities. In
this section, we discuss some observations made during the study,
threats to the validity of the study, comments made by the participants
during the study, modifications to the original algorithm to help
generalize it for more repositories, and future work.

@@

\dmg{I think you mean challenge using gitk, right? so be explicit}

During the course of the study, we noticed a consistent theme among the
participants. The most difficult challenge was identifying the master
branch of the repository. Given that single-commit merge trees are the
most common tree, and that they are relatively simple, we were
interested in seeing if the participants would be able to identify the
tree easily, but the results suggest that this was not the case.
\dmg{the following is hard to understand. even i don't quite get what you mean}
The
issue primarily stemmed from not being able to identify the master
branch. Most of the participants realized that they should not be going
backward, toward the initial commit, very early on. Furthermore, many
would identify the correct commit; however, they would start listing
every merge along the master branch as being a merge in the tree, and
including the commits from that merge. In the case of the single commit,
simply identifying the master branch would most likely be sufficient in
providing a clearer understanding of how a commit or flat tree is
integrated into the master branch of the repository.

While, in general, the participants of the study preferred using \tool
to Gitk, there were some noticeable usability issues. The most prominent
issue being the navigation within the tree. The participants expected
that the tabs would update when they clicked on a node in the tree,
instead of having to click the link to the commit after clicking the
node. This was true in both the Reingold-Tilford tree and the Pack tree.
The second major usability issue was the inconsistency between the list
tree and the other tree visualizations. The list tree is rooted at the
repository event that the user is current looking at, while the other
tree visualizations show the entire tree, and use bright orange to
highlight the current node.

@@

While git includes gitk and the git log provides a view of the DAG with
\verb|git log --graph|, many of the participants in our study were
unfamiliar with the DAG visualizations shown by these tools. A possible
reason for this may be that unless a repository reaches a sufficient
complexity, both in branching and number of commits, these tools are
necessary for the normal operation of git. Furthermore, fear of the
complexity caused by branching may cause personal and academic
repositories to have relatively simple structures. While they were
unfamiliar with the DAG structure, these users were also unfamiliar with
the \mt model, and were still able to better understand and summarize
the repository using \tool.

\dmg{this paragraph i don't get:}
Each question was evaluated independently of the results from the
previous question in the summarization tasks. Had we modified the
correct answer to fit the conceptual understanding of which commits were
included in the tree, we may have witnessed different results. A
modification to the study may include providing the participants with
the correct set of commits prior to the summarization tasks, after the
merge task set.

>>>>

% vim:set et sw=2 ts=4 tw=72:
% Jun 24, 2017

The results show conclusive results; the DAG is unable to adequately
provide users with an explanation of the changes being made in the
repository, the DAG is not easily understood, and the DAG is unable to
provide summarizations of the changes being made at a merge. Meanwhile,
the \mt model is able to provide a better explanation of how a commit is
integrated, along with better summarization capabilities. In this
section, we discuss some observations made during the study, threats to
the validity of the study, comments made by the participants during the
study, modifications to the original algorithm to help generalize it for
more repositories, and future work.

@@

% \dmg{I think you mean challenge using gitk, right? so be explicit}

During the course of the study, we noticed an issue that was consistent
among the participants when using Gitk; the most difficult challenge was
identifying the master branch of the repository. Many participants would
identify the tag as being the ``master branch'', but a tag is not a
branch. The visualization of the DAG does not provide any notion of
which branch being shown is the master branch. The default assumption is
that the first branch on the left is the master branch, but this is not
a good assumption. Furthermore, the visualization in Gitk is not
consistent, the ordering of the branches changes between runs, so
identifying the branch once will not guarantee that it will be
identifiable after restarting Gitk.

If users were able to easily identify the master branch using the
visualizations of the DAG, the results of the study would likely be very
different. This would certainly be the case for the single-commit tree;
it is trivial to summarize information about a single item, but
identifying that there was only a single commit in the tree was
non-trivial. This information is important because many trees contain a
single commit, and the structure of these trees is identical to the
structure of flat trees. Flat trees are the most common branching
structure in the repository, with the root being the parent to all of
the other nodes in the tree. To improve summarization and comprehension,
it would likely by sufficient to indicate the master branch in the DAG.

While, in general, the participants of the study preferred using \tool
to Gitk, there were some noticeable usability issues. The most prominent
issue being the navigation within the tree. The participants expected
that the tabs would update when they clicked on a node in the tree,
instead of having to click the link to the commit after clicking the
node. This was true in both the \rt tree and the Pack tree. The second
major usability issue was the inconsistency between the list tree and
the other tree visualizations. The list tree is rooted at the repository
event that the user is current looking at, while the other tree
visualizations show the entire tree, and use bright orange to highlight
the current node.

@@

\textbf{Internal Threats:}

The repository contains many commits, of which we only chose two. The
commits chosen were randomly chosen to be representative of those in the
repository. Additional commits could be added, but this would include
additional time costs.

The answers provided by the participants to earlier tasks were not
taken into account for determining the accuracy. If the participant was
unable to determine the correct commits for the tree within the
conceptual task group, the summarization results would be recorded as
wrong, even if the summarization they provided was correct for the set
of commits selected to be part of the tree. A further study could be
done where the correct answer was provided to the participant after each
task so that errors would not propagate between tasks.

To mitigate the effects of order-bias, the order that the tasks were
performed was shuffled between participants. Furthermore, the order that
the tool the tasks were performed on changed between participants, and
the order of the commits that the tasks were performed on was changed
between participants.

We were only able to work with 12 participants. We apply various
statistical tests to ensure that our findings are reasonable.

\textbf{External Threats:}

While git includes gitk and git log provides a view of the DAG with
\verb|git log --graph|, many of the participants in our study were
unfamiliar with the DAG visualizations shown by these tools. Other tools
may provide better summarizations and visualizations, but we are unaware
of them. We investigated the gui tools for Linux on the git
website\footnote{\url{https://git-scm.com/download/gui/linux}}, but none
of the tools, with the exception of Gitk and git log, were able to
produce a visualization of the DAG for the Linux repository.
Investigating the tools more deeply, it appears that GitKraken is the
most popularly cited git client. The visualization provided by GitKraken
was centered around the same DAG metaphor used by git log. GitKraken was
unable to produce a visualization for the Linux repository on our
system.

% A possible
% reason for this may be that unless a repository reaches a sufficient
% complexity, both in branching and number of commits, these tools are
% unnecessary for the normal operation of git. Furthermore, fear of the
% complexity caused by branching may cause personal and academic
% repositories to have relatively simple structures. While they were
% unfamiliar with the DAG structure, these users were also unfamiliar with
% the \mt model, and were still able to better understand and summarize
% the repository using \tool.
